---
title: "Prepare Phyloseq"
author: "Bioinformatics Core"
output:
    html_document:
      keep_md: true
---

### Create a new RStudio project

Open RStudio and create a new project, for more info see <https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects>

* File > New Project > New Directory > New Project (name the new directory, Ex. Differential_Expression) and check "use packrat with this project" if present and you want to use packrat.

Learn more about packrat see <https://rstudio.github.io/packrat/>

We first need to make sure we have the necessary packages, phyloseq, ggplot2, gridExtra, gridR, ape, and edgeR are installed (if not install it), and then load

In the R console run the following commands
```{r setup, results='hide', warning=FALSE, error=FALSE, message=FALSE, collapse=TRUE}

if (!any(rownames(installed.packages()) == "phyloseq")){
  if (!requireNamespace("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
  BiocManager::install("phyloseq")
}
library(phyloseq)

if (!any(rownames(installed.packages()) == "biomformat")){
  if (!requireNamespace("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
  BiocManager::install("biomformat")
}
library(biomformat)

if (!any(rownames(installed.packages()) == "ggplot2")){
  if (!requireNamespace("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
  BiocManager::install("ggplot2")
}
library(ggplot2)

if (!any(rownames(installed.packages()) == "gridExtra")){
  if (!requireNamespace("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
  BiocManager::install("gridExtra")
}
library(gridExtra)

if (!any(rownames(installed.packages()) == "vegan")){
  if (!requireNamespace("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
  BiocManager::install("vegan")
}
library(vegan)

if (!any(rownames(installed.packages()) == "edgeR")){
  if (!requireNamespace("BiocManager", quietly = TRUE))
      install.packages("BiocManager")
  BiocManager::install("edgeR")
}
library(edgeR)
```

### Download the template Markdown workshop document and open it

In the R console run the following command
```{r download2, results='hide', warning=F, error=F, message=FALSE}
download.file("https://raw.githubusercontent.com/ucdavis-bioinformatics-training/2019_September_UCD_Microbial_Community_Analysis_Workshop/master/MCA_Workshop_R/phyloseq.Rmd", "MCA_phyloseq.Rmd")
```

### Download the data file for the workshop document and preview/open it

```{r download2, results='hide', warning=F, error=F}
download.file("https://raw.githubusercontent.com/ucdavis-bioinformatics-training/2019_September_UCD_Microbial_Community_Analysis_Workshop/master/MCA_Workshop_R/16sV3V5.biom", "16sV3V5.biom")
```

### Edit the file YAML portion

The top YAML (YAML ain't markup language) portion of the doc tells RStudio how to parse the document.

<pre><code>---
title: "Microbial Community Analysis in R"
author: your_name
date: current_date
output:
    html_notebook: default
    html_document: default
---</code></pre>
